Reanalysis Step 5: Outcome Analysis with Full Four-Treatment Interaction

### 2025-09-02 reanalysis-5-outcome_analysis-full_interaction.qmd
fix fix fix fix

# Configuration for stability checking
MAX_SE_THRESHOLD <- 10  # Maximum acceptable standard error
MIN_N_PER_CELL <- 5    # Minimum observations per interaction cell
TEST_SINGLE_IMPUTATION <- TRUE  # Test on single imputation first
```

@@

# Create consistent analysis variables across all imputed datasets
for (i in 1:n_imputations) {
  imputed_datasets[[i]] <- imputed_datasets[[i]] |>
    mutate(
      # Create 4-level treatment factor (hierarchical priority)
      # Priority: anti_VEGF > PRP > focal_laser > other_inject > none
      treatment_type = case_when(
        anti_VEGF == 1 ~ "anti_VEGF",
        PRP_flag == 1 ~ "PRP",
        focal_laser_flag == 1 ~ "focal_laser",
        other_inject == 1 ~ "other_inject",
        TRUE ~ "none"
      ),
      treatment_type = factor(treatment_type, 
                              levels = c("none", "other_inject", "focal_laser", 
                                       "PRP", "anti_VEGF")),

@@

# Check treatment distribution
treatment_dist <- table(imputed_datasets[[1]]$treatment_type)
log_message("Treatment type distribution in first imputation:")
for (tx in names(treatment_dist)) {
  log_message(paste0("  ", tx, ": ", treatment_dist[tx]))
}
```

>>>>

# Configuration for stability checking
MAX_SE_THRESHOLD <- 10  # Maximum acceptable standard error
MIN_N_PER_CELL <- 5    # Minimum observations per interaction cell
TEST_SINGLE_IMPUTATION <- TRUE  # Test on single imputation first

# Treatment categorization approach
# Option 1: "hierarchical" - assigns to highest priority treatment (focal > anti_VEGF > PRP > other)
#           NOTE: focal_laser now has highest priority to preserve all 36 cases
# Option 2: "collapsed" - combines focal_laser with other_inject due to small n
# Option 3: "first_only" - uses only the first/primary treatment
TREATMENT_APPROACH <- "hierarchical"  # Using hierarchical with focal priority
```

@@

# Create consistent analysis variables across all imputed datasets
for (i in 1:n_imputations) {
  imputed_datasets[[i]] <- imputed_datasets[[i]] |>
    mutate(
      # First, check treatment overlaps for diagnostic purposes
      n_treatments = other_inject + anti_VEGF + focal_laser_flag + PRP_flag,

      # Create treatment factor based on chosen approach
      treatment_type = case_when(
        TREATMENT_APPROACH == "hierarchical" ~ case_when(
          focal_laser_flag == 1 ~ "focal_laser",  # Highest priority to preserve all 36 cases
          anti_VEGF == 1 ~ "anti_VEGF",
          PRP_flag == 1 ~ "PRP",
          other_inject == 1 ~ "other_inject",
          TRUE ~ "none"
        ),
        TREATMENT_APPROACH == "collapsed" ~ case_when(
          anti_VEGF == 1 ~ "anti_VEGF",
          PRP_flag == 1 ~ "PRP",
          focal_laser_flag == 1 | other_inject == 1 ~ "other_or_focal",
          TRUE ~ "none"
        ),
        TRUE ~ "none"  # fallback
      ),

      # Set factor levels based on approach
      treatment_type = if (TREATMENT_APPROACH == "collapsed") {
        factor(treatment_type, 
               levels = c("none", "other_or_focal", "PRP", "anti_VEGF"))
      } else {
        factor(treatment_type,
               levels = c("none", "other_inject", "focal_laser", "PRP", "anti_VEGF"))
      },

@@

# Check treatment distribution and overlaps
first_data <- imputed_datasets[[1]]

log_message(paste0("Using treatment approach: ", TREATMENT_APPROACH))
if (TREATMENT_APPROACH == "collapsed") {
  log_message("Combining focal_laser and other_inject into 'other_or_focal' category")
}

# Check how many patients have multiple treatments
multi_treatment <- table(first_data$n_treatments)
log_message("Number of treatments per patient:")
for (n in names(multi_treatment)) {
  log_message(paste0("  ", n, " treatment(s): ", multi_treatment[n]))
}

# Check specific overlaps with focal laser
if (sum(first_data$focal_laser_flag) > 0) {
  focal_patients <- first_data[first_data$focal_laser_flag == 1, ]
  log_message(paste0("Total focal laser patients: ", nrow(focal_patients)))
  log_message(paste0("  Focal only: ", sum(focal_patients$n_treatments == 1)))
  log_message(paste0("  Focal + anti_VEGF: ", sum(focal_patients$anti_VEGF == 1)))
  log_message(paste0("  Focal + PRP: ", sum(focal_patients$PRP_flag == 1)))
  log_message(paste0("  Focal + other: ", sum(focal_patients$other_inject == 1)))
}

treatment_dist <- table(first_data$treatment_type)
log_message("Treatment type distribution (hierarchical assignment):")
for (tx in names(treatment_dist)) {
  log_message(paste0("  ", tx, ": ", treatment_dist[tx]))
}

# Create a cross-tab to show the hierarchical effect
focal_assignment <- table(
  focal_laser = first_data$focal_laser_flag,
  assigned_treatment = first_data$treatment_type
)
log_message("Where focal laser patients were assigned:")
print(focal_assignment)
```
